Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massively Distributed Evolutionary Computation

### 2015-07-27 volunteer-clouds-ieee-ci-si.tex
Adding global computing desirable features

The web used as a resource for distributed computing has a
long history, and almost since the beginning it was used for 
evolutionary computation even before JavaScript was
introduced as a browser-based scripting language in 1997. Before that
time other technologies,
%other technologies?
% changed - JJ
 including Flash animations and VBScript, ActiveX or Java applets were
used; for instance, Java was pointed out in \cite{soares1998get} as a ``language for
internet'', providing some advantages such as multi-architecture compatibility or 
security mechanisms and was proposed in \cite{chandy1996world} as the
basis for the creation of a world-wide peer to peer system using Java
compiled programs. The ATLAS system
\cite{Baldeschwieler:1996:TIG:504450.504482} also used Java together
with another language called Cilk; all these papers indicated that
the introduction of an architecture-independent system such as the
Java virtual machine opened the doors for ultrascale, Internet-wide,
metacomputers.

Using any of these languages allowed the creation
of client-server applications in the browser, provided that the user had
installed the required virtual machine plugin first, but with the advantage of
providing a single and universal way of accessing resources via an
URI, something that is not guaranteed using desktop applications. However, this style of
computing presents several issues and challenges which were addressed
as early as 1996 by Baratloo et al. in their system Charlotte:
\cite{baratloo1996charlotte}. This system, which implemented a
distributed shared memory, was inspired by previous work
on parallel computing systems PVM and MPI, but solved most issues by
using the Java virtual machine embedded in the browser,
which provides
security to the user, uniformity of computing environment
independently of the operating system and  access to a common data store in the
server. Other issues such as scheduling are solved on the server. This
paper was later extended and published in a journal in 1999
\cite{baratloo1999charlotte}, but it is interesting to note that
already in the early stages of the web the possibility of
browser-based distributed computing was already researched.

>>>>

The web used as a resource for distributed computing has a
long history, and almost since the beginning it was used for 
evolutionary computation even before JavaScript was
introduced as a browser-based scripting language in 1997. Before that
time other technologies,
%other technologies?
% changed - JJ
 including Flash animations and VBScript, ActiveX or Java applets were
used; for instance, Java was pointed out in \cite{soares1998get} as a ``language for
internet'', providing some advantages such as multi-architecture compatibility or 
security mechanisms and was proposed in \cite{chandy1996world} as the
basis for the creation of a world-wide peer to peer system using Java
compiled programs. The ATLAS system
\cite{Baldeschwieler:1996:TIG:504450.504482} also used Java together
with another language called Cilk; all these papers indicated that
the introduction of an architecture-independent system such as the
Java virtual machine opened the doors for ultrascale, Internet-wide,
metacomputers. This paper also establishes the desirable features of
any {\em global computing} infrastructure, that can also be applied to
distributed evolutionary computation experiments that rely in
volunteer computing: Scalability, heterogeneity, fault tolerance,
adaptive parallelism, safety, anonymity, hierarchy, ease of use and
reasonable performance. Of all these, probably {\em hierarchy}, that
is, the capability of using preferentially resources in the same
organization could be either disregarded or included in adaptivity,
leaving us with eight desirable properties.

Fulfilling these properties could be achieved using any of the
languages, and associated technologies, mentioned above, provided that the user had
installed the required virtual machine plugin first, but with the advantage of
providing a single and universal way of accessing resources via an
URI, something that is not guaranteed using desktop applications. However, this style of
computing presents several issues and challenges which were addressed
as early as 1996 by Baratloo et al. in their system Charlotte:
\cite{baratloo1996charlotte}. This system, which implemented a
distributed shared memory, was inspired by previous work
on parallel computing systems PVM and MPI, but solved most issues by
using the Java virtual machine embedded in the browser,
which provides
security to the user, uniformity of computing environment
independently of the operating system and  access to a common data store in the
server. Other issues such as scheduling are solved on the server. This
paper was later extended and published in a journal in 1999
\cite{baratloo1999charlotte}, but it is interesting to note that
already in the early stages of the web the possibility of
browser-based distributed computing was already researched.
